Informing forest carbon inventories under the Paris Agreement using the Global Forest Carbon Database (ForC v4.0)

### 2024-02-02 doc/manuscript/ForC-EFDB manuscript.Rmd
revising manuscript

```{r eval = TRUE, echo=FALSE, warning=FALSE, results='hide'}
#Tables
table_pools =1

table_recommendations = 3

@@

#Cited sections (listed here because these can change)
section_C_pools = 2.1 
section_land_use_categories = 2.2
section_new_variables = 3.2
section_C_variable_mapping = "4.1.1"
section_land_use_mapping = "4.1.2"
section_database_needs = "6.1"
section_new_data_needs = "6.2"
section_data_reporting_needs = "6.3"

#Appendices
appendix_ForC_updates = "Methods S1"
appendix_IPCC_definitions = "Notes S1"
table_ForCchanges = "Table S1"
table_ForCfieldmapping = "Table S2"
table_variables ="S3"

@@

<!-- (1) the research conducted or aims, including the rationale-->
- Humans have been influencing Earth's climate via transformative impacts on forests for >12,000 years, and forests are now recognized as critical to climate change mitigation under the Paris Agreement.
However, the efficacy of climate change mitigation planning and reporting depends on having quality data on forest carbon (C) stocks and changes.
The Emission Factor Database (EFDB) of the International Panel on Climate Change (IPCC) is intended to be a definitive source for such data, but needs comprehensive and well-documented data to be so.
<!-- (2) methods-->
- To facilitate submission of forest C estimates from scientific studies to EFDB, we develop and document a process for semi-automated data submission from the Global Forest C database (ForC v4.0), which is the largest compilation of ground-based forest C estimates. 
We then assess the data currently available through ForC and provide recommendations for improving forest data collection, analysis, and reporting.
<!-- (3) key results-->
- As of `r format(Sys.time(), '%B %Y')`, ForC contained ~`r n_ind_records_EFDB_variables` independent records relevant to EFDB, `r n_records_posted_EFDB` of which had been submitted and posted to EFDB.
Records were unevenly distributed across variables (skewed towards aboveground biomass stocks) and geographic regions (skewed towards temperate forests).
`r round((n_records_reviewed_or_added-n_records_sent_to_EFDB)/n_records_sent_to_EFDB,2)*100`% of ForC records reviewed could not be submitted because the original publication lacked required information.
<!-- (4) the main conclusion, including the key points of discussion-->
- In the future, ground-based forest C estimates should target gaps in the record, and studies should ensure that they report all information necessary for inclusion in EFDB. 
Given that climate change is rapidly impacting the world’s forests, timely reporting of recent estimates will be especially critical to accurate forest C inventories.

@@

@@

- helps bridge scientific research accounting under Paris agreement

@@

As of `r format(Sys.time(), '%B %d, %Y')`<!--inserts date upon knit-->, ForC (v4.0) contained `r n_independent_records_ForC` independent records (`r n_records_ForC` total), `r n_ind_records_EFDB_variables` of which were for the `r n_relevant_variables` variables relevant to EFDB (Fig. `r fig_variable_mapping`).
We had reviewed or added `r n_records_reviewed_or_added` records for submission to EFDB, `r n_records_sent_to_EFDB` of which met the criteria for submission, and `r n_records_posted_EFDB` of which had been reviewed by EFDB's review panel, accepted, and posted (Table `r table_variables`).
The `r round((n_records_reviewed_or_added -n_records_sent_to_EFDB )/n_records_reviewed_or_added,2)*100`% attenuation between records we reviewed and those submitted to EFDB was attributable to the presence of digitized records and records where a variable's value had been calculated as the sum or difference of related variables rather than presented directly in the text.
The discrepancy between the number of records sent and that posted to EFDB is because some records (`r round((1-(n_records_posted_EFDB/n_records_sent_to_EFDB)),2)*100`%) were deemed not applicable to EFDB by the review panel.
As of `r format(Sys.time(), '%B %d, %Y')`, the `r n_records_sent_to_EFDB` ForC records posted in EFDB represented `r percent_forest_records`% of the total EFDB records for forest land.

ForC v4.0 contained records for `r n_relevant_variables_with_data` of the `r n_relevant_variables` variables (or closely-related variable groups) relevant to EFDB (Fig. `r fig_variable_mapping`, Table `r table_variables`).
Records were submitted for `r n_relevant_variables_records_sent` variables (or closely-related variable groups), including variables from each C pool (Table `r table_variables`).
The records available in ForC v4.0 -- and the subset submitted to EFDB -- were very unevenly distributed across variables (Table `r table_variables`). 
The majority (`r round(n_records_sent_stocks/n_records_sent_to_EFDB, 2)*100`%) of records submitted were for C stocks, including
`r round(n_records_sent_biomass/n_records_sent_to_EFDB,2)*100`% for total biomass, 
`r round(n_records_sent_agb/n_records_sent_to_EFDB,2)*100`% for aboveground biomass,
`r round(n_records_sent_agb_components/n_records_sent_to_EFDB,2)*100`% for components of aboveground biomass (wood or foliage), 
`r round(n_records_sent_rootbiomass/n_records_sent_to_EFDB,2)*100`% for root biomass, 
`r round(n_records_sent_rootbiomass_components/n_records_sent_to_EFDB,2)*100`% for components of root biomass (coarse or fine roots), 
`r round(n_records_sent_deadwood/n_records_sent_to_EFDB,2)*100`% for dead wood, 
`r round(n_records_sent_deadwood_components/n_records_sent_to_EFDB,2)*100`% for components of dead wood (standing or fallen), 
`r round(n_records_sent_litter/n_records_sent_to_EFDB,2)*100`% for litter (entire O horizon or OL layer component),
and `r round(n_records_sent_soil/n_records_sent_to_EFDB,2)*100`% for SOM/ SOC.
Increment records totaled `r round(n_records_sent_increments/n_records_sent_to_EFDB*100, 0)`% of records submitted, all but four of which described aboveground biomass.
The remaining `r round(n_records_sent_fluxes/n_records_sent_to_EFDB*100, 0)`% of records submitted described fluxes, all of which were either inputs or outputs to the aboveground biomass pool, a subset of which also described inputs to the dead wood or litter pool (Fig. `r fig_variable_mapping`,Table `r table_variables`).

@@

** RECOMENDATIONS **

Based on our experience contributing forest C data to EFDB via ForC, we make several recommendations as to how scientists can improve forest C records in EFDB through database work (section `r section_database_needs`), new data collection and analysis (section `r section_new_data_needs`), and reporting (section `r section_data_reporting_needs`).

## Database needs {-}

There is vast potential to expand forest C data in EFDB by completing the process of reviewing and submitting data that are already in ForC (Figs. `r fig_map`-`r fig_histograms`).
So far, only ~`r round(n_records_sent_to_EFDB/n_ind_records_EFDB_variables,2)*100`% of the EFDB-relevant data in ForC have been submitted to EFDB.
Although this process requires manual review of records, the submission of new records to EFDB is hugely facilitated by the fact that most pertinent information for each record is already entered in ForC and can be easily prepared for submission to EFDB using the system developed here.
Future efforts to review studies for submission should optimize for representation across geographic regions, forest types, and variables, giving priority to those from currently under-represented regions and forest types (Figs. `r fig_map`-`r fig_histograms`, Table `r table_variables`), to records from countries relying on existing data for their greenhouse gas inventories (Tier 1 or 2 methodology), to the variables most needed by EFDB users, and to the more contemporary records.

@@

## Data collection and analysis needs {-}

New data collection and analysis is needed to fill notable knowledge gaps.
While aboveground biomass stocks in particular have received -- and continue to receive -- by far the most research attention [Table `r table_variables`, @anderson-teixeira_carbon_2021;@dubayah_global_2020; @quegan_european_2019; @nisar_nasaisro_2018], production of an accurate global map of forest C stocks remains an ongoing challenge [@araza_decade_2023]. 
Other pools and variables remain very poorly quantified [Table `r table_variables`, @anderson-teixeira_carbon_2021], introducing substantive uncertainties into global forest C budgets [@pan_large_2011;@harris_global_2021].
Furthermore, data distribution is uneven across forest types and geographical regions (Figs. `r fig_map`-`r fig_histograms`).
For instance, data on C cycling of tropical forests -- particularly in Africa -- remains relatively sparse, in large part due to substantial barriers to data collection and distribution [@delima_making_2022].
Significant investment in research and researchers focused on ground-based measurement of forest C in such regions will be important to filling knowledge gaps in forest C cycling [@delima_making_2022;@araza_decade_2023; @labrière_forest_2023].

Several EFDB-relevant variables have not been calculated and presented as frequently as would be possible given existing forest census data and minimal extra research effort.
For example, aboveground woody mortality (*woody.mortality_ag*) and aboveground biomass increment (*delta.agb*) can be calculated from the same census data as aboveground woody productivity (*ANPP_woody*), yet the latter has received far more research attention, and correspondingly has far more records in ForC [Table `r table_variables`, @anderson-teixeira_carbon_2021; but see @piponiot_distribution_2022]. 
Similarly, live coarse root biomass, total biomass, and changes in both of these pools could in theory be estimated in parallel with aboveground biomass, with the greatest barrier being that allometric models for estimating root biomass are not as reliable or easily available as are those for aboveground biomass [@chave_improved_2014;@rejou-mechain_biomass_2017;@gonzalez-akre_allodb_2022].
However, while equations for estimating root (and thereby total) biomass require improvement, they do exist for many forest types [e.g., @brassard_coarse_2011; @chojnacky_updated_2014;  @waring_overlooking_2017; @mokany_critical_2006].
In addition, standing dead trees are captured in most forest censuses and could be used to estimate standing dead wood, although additional data on breakage would be needed for accurate estimates. 
We recommend that, when possible, researchers calculate and report these variables, following the reporting guidelines specified in section `r section_data_reporting_needs`.

Filling knowledge gaps in other EFDB-relevant variables will require more effort, but this effort is warranted given their importance for estimating forest C stock chnages.
Although aboveground biomass is the most studied variable considered here (Table `r table_variables`) and is the target of satellite missions [@dubayah_global_2020; @quegan_european_2019; @nisar_nasaisro_2018], significant ground-based research effort is required to create accurate global maps of forest biomass and changes therein [@duncanson_importance_2019; @labrière_forest_2023;@calders_laser_2022].
Given observations of increasing tree mortality in some forested regions [@mcdowell_pervasive_2020], better characterization of forest dead wood will be critical.
Additionally, C stocks in forest organic horizons and soils can be quite substantial and highly uncertain in many parts of the world [@tifafi_large_2018].
Significant investment in ground-based forest research will be critical to filling these gaps.

## Data reporting needs {-}

@@

For those compiling published records (e.g., for meta-analyses), the data set can have added value if all information required by EFDB is extracted from original publications. 
This includes -- but is not limited to -- retaining original values as presented without modification or rounding, noting whether data were digitized, recording confidence intervals, and recording all required fields (as indicated in the EFDB's bulk import template).
The significant effort required to map a database into EFDB has been accomplished here (`r table_ForCfieldmapping`), and we welcome other researchers to use the ForC template.

Once EFDB-relevant data are available in peer-reviewed publications, they may be submitted directly to EFDB or may use the ForC - EFDB data pipeline developed here.
For individual publications, the former option will generally be more efficient.
However, data incorporated into ForC as well as EFDB will be more broadly useful; for example, these data may be used for basic science [e.g., @banburymorgan_global_2021; @anderson-teixeira_carbon_2021], analyses of forest-based climate change mitigation potential [e.g., @cook-patton_mapping_2020; @goldstein_protecting_2020], and model benchmarking [@fer_ecosystem_2021].

@@

The ForC database contains large numbers of records that could potentially be useful for estimating C stock changes applying methodological guidance provided by the IPCC.
Here we have developed a framework for submitting these records to the EFDB, thus making those data more accessible for reporting CO~2~ emissions and removals from forest land consistent with good practice in the IPCC guidelines  [@ipcc_2006_2006;@ipcc_2019_2019].
As of `r format(Sys.time(), '%B %d, %Y')`, we have submitted `r n_records_sent_to_EFDB` records to EFDB.
Although this represents just `r round(n_records_sent_to_EFDB/n_ind_records_EFDB_variables*100,0)`% of relevant records in ForC, it substantially increases the number of forest land records in EFDB.
The records submitted to EFDB and present in ForC are very unevenly distributed across variables, regions, and forest types (Figs. `r fig_map`-`r fig_histograms`, Table `r table_variables`), reflecting broader patterns in allocation of research effort.

Going forward, forest researchers can make their research more useful for forest C inventories under IPCC guidelines by calculating and reporting results in ways that are consistent with methodologies provided in the IPCC guidelines (Tables `r table_pools`, `r table_recommendations`).
In addition, substantial investments in research and researchers focused on ground-based measurement of forest C will be required to fill knowledge gaps and thereby increase the accuracy of forest CO~2~ inventories for forest lands under teh Paris Agreement. 
This challenge is heightened by the fact that forests are changing rapidly [e.g., @mcdowell_pervasive_2020], and data collected a decade or more in the past may no longer be relevant.
This heightens the need for an efficient system of making forest C data accessible for national greenhouse gas inventories.
We view the system developed here for submitting ForC data to the IPCC EFDB as one important step towards that goal.

@@

>>>>

```{r eval = TRUE, echo=FALSE, warning=FALSE, results='hide'}
#Tables
table_pools =1
table_recommendations = 2

@@

#Appendices
appendix_ForC_updates = "Methods S1"
appendix_IPCC_definitions = "Notes S1"
table_ForCchanges = "Table S1"
table_ForCfieldmapping = "Table S2"
table_variables ="Table S3"

@@

<!-- (1) the research conducted or aims, including the rationale-->
- Humans have been influencing Earth's climate via transformative impacts on forests for >12,000 years, and forests are now recognized as critical to climate change mitigation under the Paris Agreement.
However, the efficacy of climate change mitigation planning and reporting depends on having quality data on forest carbon (C) stocks and changes.
The Emission Factor Database (EFDB) of the International Panel on Climate Change (IPCC) is intended to be a definitive source for such data, but needs comprehensive and well-documented data to be so.
<!-- (2) methods-->
- To facilitate submission of forest C estimates from scientific studies to EFDB, we develop and document a process for semi-automated data submission from the Global Forest C database (ForC v4.0), which is the largest compilation of ground-based forest C estimates. 
We then assess the data currently available through ForC and provide recommendations for improving forest data collection, analysis, and reporting.
<!-- (3) key results-->
- As of `r format(Sys.time(), '%B %Y')`, ForC contained ~`r n_ind_records_EFDB_variables` independent records relevant to EFDB, `r n_records_posted_EFDB` of which had been submitted and posted to EFDB.
Records were unevenly distributed across variables (skewed towards aboveground biomass stocks) and geographic regions (skewed towards temperate forests).
`r round((n_records_reviewed_or_added-n_records_sent_to_EFDB)/n_records_reviewed_or_added,2)*100`% of ForC records reviewed could not be submitted because the original publication lacked required information.
<!-- (4) the main conclusion, including the key points of discussion-->
- In the future, ground-based forest C estimates should target gaps in the record, and studies should ensure that they report all information necessary for inclusion in EFDB. 
Given that climate change is rapidly impacting the world’s forests, timely reporting of recent estimates will be especially critical to accurate forest C inventories.

@@

climate change, database, forest carbon, greenhouse gas inventory, International Panel on Climate Change (IPCC), natural climate solutions

@@

Human interactions with forests have shaped Earth's climate for >12,000 years and will continue to do so as we target net-zero emissions.
Accurately characterizing these climate impacts requires accurate forest carbon data available for national greenhouse gas inventories conducted under the Paris Agreement. 
Here we develop a semi-automated process for submitting forest carbon measurements from the largest relevant scientific database to the International Panel on Climate Change's Emission Factor Database, which currently has sparse forest carbon data.
Building this bridge from scientific research to international policy is an important step towards managing forests in a net-zero motivated future.

@@

As of `r format(Sys.time(), '%B %d, %Y')`<!--inserts date upon knit-->, ForC (v4.0) contained `r n_independent_records_ForC` independent records (`r n_records_ForC` total), `r n_ind_records_EFDB_variables` of which were for the `r n_relevant_variables` variables relevant to EFDB (Fig. `r fig_variable_mapping`).
We had reviewed or added `r n_records_reviewed_or_added` records for submission to EFDB, `r n_records_sent_to_EFDB` of which met the criteria for submission, and `r n_records_posted_EFDB` of which had been reviewed by EFDB's review panel, accepted, and posted (`r table_variables`).
The `r round((n_records_reviewed_or_added -n_records_sent_to_EFDB )/n_records_reviewed_or_added,2)*100`% attenuation between records we reviewed and those submitted to EFDB was attributable to the presence of digitized records and records where a variable's value had been calculated as the sum or difference of related variables rather than presented directly in the text.
The discrepancy between the number of records sent and that posted to EFDB is because some records (`r round((1-(n_records_posted_EFDB/n_records_sent_to_EFDB)),2)*100`%) were deemed not applicable to EFDB by the review panel.
As of `r format(Sys.time(), '%B %d, %Y')`, the `r n_records_sent_to_EFDB` ForC records posted in EFDB represented `r percent_forest_records`% of the total EFDB records for forest land.

ForC v4.0 contained records for `r n_relevant_variables_with_data` of the `r n_relevant_variables` variables (or closely-related variable groups) relevant to EFDB (Fig. `r fig_variable_mapping`,  `r table_variables`).
Records were submitted for `r n_relevant_variables_records_sent` variables (or closely-related variable groups), including variables from each C pool (`r table_variables`).
The records available in ForC v4.0 -- and the subset submitted to EFDB -- were very unevenly distributed across variables (`r table_variables`). 
The majority (`r round(n_records_sent_stocks/n_records_sent_to_EFDB, 2)*100`%) of records submitted were for C stocks, including
`r round(n_records_sent_biomass/n_records_sent_to_EFDB,2)*100`% for total biomass, 
`r round(n_records_sent_agb/n_records_sent_to_EFDB,2)*100`% for aboveground biomass,
`r round(n_records_sent_agb_components/n_records_sent_to_EFDB,2)*100`% for components of aboveground biomass (wood or foliage), 
`r round(n_records_sent_rootbiomass/n_records_sent_to_EFDB,2)*100`% for root biomass, 
`r round(n_records_sent_rootbiomass_components/n_records_sent_to_EFDB,2)*100`% for components of root biomass (coarse or fine roots), 
`r round(n_records_sent_deadwood/n_records_sent_to_EFDB,2)*100`% for dead wood, 
`r round(n_records_sent_deadwood_components/n_records_sent_to_EFDB,2)*100`% for components of dead wood (standing or fallen), 
`r round(n_records_sent_litter/n_records_sent_to_EFDB,2)*100`% for litter (entire O horizon or OL layer component),
and `r round(n_records_sent_soil/n_records_sent_to_EFDB,2)*100`% for SOM/ SOC.
Increment records totaled `r round(n_records_sent_increments/n_records_sent_to_EFDB*100, 0)`% of records submitted, all but four of which described aboveground biomass.
The remaining `r round(n_records_sent_fluxes/n_records_sent_to_EFDB*100, 0)`% of records submitted described fluxes, all of which were either inputs or outputs to the aboveground biomass pool, a subset of which also described inputs to the dead wood or litter pool (Fig. `r fig_variable_mapping`,`r table_variables`).

@@

Here we have developed a framework for submitting records from ForC v4.0, the largest compilation of ground-based forest C estimates, to the EFDB, thus making those data more accessible for reporting CO~2~ emissions and removals from forest land consistent with good practice in the IPCC guidelines  [@ipcc_2006_2006;@ipcc_2019_2019].
As of `r format(Sys.time(), '%B %d, %Y')`, `r n_records_posted_EFDB` ForC records have been posted to EFDB, which represents just `r round(n_records_posted_EFDB/n_ind_records_EFDB_variables*100,0)`% of potentially relevant records in ForC v4.0, but currently comprises `r percent_forest_records`% of the total EFDB records for forest land.
The records present in ForC and submitted to EFDB are very unevenly distributed across variables, regions, and forest types (Figs. `r fig_map`-`r fig_histograms`, `r table_variables`), reflecting broader patterns in allocation of research effort and pointing to scientific research and reporting needs.

Based on our experience contributing forest C data to EFDB via ForC, we make several recommendations as to how scientists can improve forest C records in EFDB through database work, new data collection and analysis, and reporting.

### Database needs {-}

There is vast potential to expand forest C data in EFDB by completing the process of reviewing and submitting data that are already in ForC (Figs. `r fig_map`-`r fig_histograms`).
So far, only ~`r round(n_records_sent_to_EFDB/n_ind_records_EFDB_variables,2)*100`% of the EFDB-relevant data in ForC have been submitted to EFDB.
Although this process requires manual review of records, the submission of new records to EFDB is hugely facilitated by the fact that most pertinent information for each record is already entered in ForC and can be easily prepared for submission to EFDB using the system developed here.
Future efforts to review studies for submission should optimize for representation across geographic regions, forest types, and variables, giving priority to those from currently under-represented regions and forest types (Figs. `r fig_map`-`r fig_histograms`, `r table_variables`), to records from countries relying on existing data for their greenhouse gas inventories (Tier 1 or 2 methodology), to the variables most needed by EFDB users, and to the more contemporary records.

@@

### Data collection and analysis needs {-}

New data collection and analyses are needed to fill notable knowledge gaps.
While aboveground biomass stocks in particular have received -- and continue to receive -- by far the most research attention [`r table_variables`, @anderson-teixeira_carbon_2021;@dubayah_global_2020; @quegan_european_2019; @nisar_nasaisro_2018], production of an accurate global map of forest C stocks remains an ongoing challenge [@araza_decade_2023]. 
Other pools and variables remain very poorly quantified and highly uncertain for many parts of the world [`r table_variables`,@tifafi_large_2018; @anderson-teixeira_carbon_2021], introducing substantive uncertainties into global forest C budgets [@pan_large_2011;@harris_global_2021].
Furthermore, data distribution is very uneven across forest types and geographical regions (Figs. `r fig_map`-`r fig_histograms`).
For instance, data on C cycling of tropical forests -- particularly in Africa -- remains relatively sparse, in large part due to substantial barriers to data collection and distribution [@delima_making_2022].
Significant investment in research and researchers focused on ground-based measurement of forest C in such regions will be important to filling knowledge gaps in forest C cycling [@delima_making_2022;@araza_decade_2023; @labrière_forest_2023].

Several EFDB-relevant variables have not been calculated and presented as frequently as would be possible given existing forest census data and minimal extra research effort.
For example, aboveground woody mortality and aboveground biomass increment can be calculated from the same census data as aboveground woody productivity, yet the latter has received far more research attention [`r table_variables`, @anderson-teixeira_carbon_2021]. 
Similarly, live coarse root biomass, total biomass, and changes therein could in theory be estimated in parallel with aboveground biomass, with the greatest barrier being that allometric models for estimating root biomass are not as reliable or easily available as are those for aboveground biomass [@chave_improved_2014;@rejou-mechain_biomass_2017;@gonzalez-akre_allodb_2022].
However, while equations for estimating root (and thereby total) biomass require improvement, they do exist for many forest types [e.g., @brassard_coarse_2011; @chojnacky_updated_2014;  @waring_overlooking_2017; @mokany_critical_2006].
We recommend that, when possible, researchers calculate and report all EFDB-relevant variables.

### Data reporting needs {-}

@@

For those compiling published records (e.g., for meta-analyses), the data set can have added value if all information required by EFDB is extracted from original publications. 
This includes -- but is not limited to -- retaining original values as presented without modification or rounding, noting whether data were digitized, recording confidence intervals, and recording all required fields (as indicated in the EFDB's bulk import template).
The significant effort required to map a database into EFDB has been accomplished here (`r table_ForCfieldmapping`), and we hope that it will prove useful as a model for other efforts.

@@

As human society strives to achieve net-zero greenhouse gas emissions, forest researchers can make their research more useful for forest C inventories under IPCC guidelines by calculating and reporting results in ways that are consistent with methodologies provided in the IPCC guidelines (Tables `r table_pools`, `r table_recommendations`).
In addition, substantial investments in research and researchers focused on ground-based measurement of forest C will be required to fill knowledge gaps and thereby increase the accuracy of forest CO~2~ inventories for forest lands under teh Paris Agreement. 
This challenge is heightened by the fact that forests are changing rapidly [e.g., @mcdowell_pervasive_2020], and data collected a decade or more in the past may no longer be relevant.
This heightens the need for an efficient system of making forest C data accessible for national greenhouse gas inventories.
We view the system developed here for submitting ForC data to the IPCC EFDB as one important step towards that goal.

@@

# Supplementary Information
`r appendix_IPCC_definitions`. Primer on forest land classification and carbon pools under IPCC guidelines

`r appendix_ForC_updates`. Updates to ForC (ForC v4.0)

`r table_ForCchanges`.  Updates to ForC field implemented between releases of v3.0 and v4.0

`r table_ForCfieldmapping`. Mapping of ForC fields to EFDB

`r table_variables`. Numbers of ForC records and EFDB submissions by variable

### 2024-02-02 doc/manuscript/SI_ForCv40.Rmd
revising manuscript

```{r eval = TRUE, echo=FALSE, warning=FALSE, results='hide'}
#Tables
table_pools =1
table_variables =2
table_recommendations = 3

@@

EFDB data submissions required information that was not recorded in previous versions of ForC, but for which new fields were created for EFDB compatibility (`r table_ForCchanges`).
It was therefore necessary to return to original publications to retrieve relevant information, including 
(1) estimates in original units, 
(2) confidence intervals (when not already in ForC),
(3) whether records of interest were presented in tables or text or digitized from figures (EFDB will not accept digitized data), 
(4) whether records of interest were presented directly, as opposed to having been calculated from related variables (for example, if a study presents aboveground biomass and root biomass but not total biomass, EFDB would not accept the sum of these as a valid record of total biomass)
We also checked that existing ForC records were complete and correct.

>>>>

```{r eval = TRUE, echo=FALSE, warning=FALSE, results='hide'}
#Tables
table_pools =1
table_recommendations = 2

@@

EFDB data submissions required information that was not recorded in previous versions of ForC, but for which new fields were created for EFDB compatibility (`r table_ForCchanges`).
It was therefore necessary to return to original publications to retrieve relevant information, including 
(1) estimates in original units, 
(2) confidence intervals (when not already in ForC),
(3) whether records of interest were presented in tables or text or digitized from figures (EFDB will not accept digitized data), 
and (4) whether records of interest were presented directly, as opposed to having been calculated from related variables (for example, if a study presents aboveground biomass and root biomass but not total biomass, EFDB would not accept the sum of these as a valid record of total biomass).
We also checked that existing ForC records were complete and correct.

### 2024-02-02 doc/manuscript/SI_tables.Rmd
revising manuscript

\newpage
**`r table_variables`. Numbers of records of ForC variables (or closely related variable groups) relevant to, and sent to, EFDB.**
```{r table_variables, eval = TRUE, echo=FALSE, warning=FALSE}
C_variables_table <- read.csv("figures_tables/C_variables.csv")
names(C_variables_table) <- stringr::str_replace_all(names(C_variables_table), "\\.", " ")
C_variables_table[is.na(C_variables_table)]<- "" #replace NAs with blanks

>>>>

\newpage
**`r table_variables`. Numbers of ForC records and EFDB submissions by variable.**
```{r table_variables, eval = TRUE, echo=FALSE, warning=FALSE}
C_variables_table <- read.csv("figures_tables/C_variables.csv")
names(C_variables_table) <- stringr::str_replace_all(names(C_variables_table), "\\.", " ")
C_variables_table[is.na(C_variables_table)]<- "" #replace NAs with blanks
